Ice-shelf ocean boundary layer dynamics from large-eddy simulations

### 2021-07-12 main.tex

\documentclass[draft]{agujournal2019}
\usepackage{url} 
\usepackage{lineno}
\usepackage{graphicx}
%\usepackage{xr}
\usepackage{float}
%\externaldocument{A_suppl_fig}
\usepackage{amsmath}
\linenumbers
\draftfalse
\journalname{JGR: Oceans}

@@

\textbf{Plain language summary}\\

@@

% Motivation - SLR
The largest source of uncertainty in future sea level rise is the potential loss of ice from the Antarctic Ice Sheet. The rate of grounded ice loss is highly sensitive to the melting of ice shelves, which drain over 80\% of Antarctica’s grounded ice. The Ice-shelf/Ocean Boundary Layer (IOBL) controls ice-shelf melting by regulating oceanic heat and salt fluxes to the ice shelf base. Thus, accurate predictions of ice-shelf melting depend on capturing the turbulent dynamics of the IOBL. This representation is also critical for evaluating the sensitivity of the coupled land ice-ocean system to changes in ocean conditions. Of particular concern is the sensitivity of ice-shelf melting to increasing seawater temperature, a trend observed along a wide swath of the West Antarctic coastline and a potential trigger for West Antarctic Ice Sheet collapse.

% Shortcomings of ocean-model representations of IOBL
One indication that ocean models do not capture these dynamics is that the modeled thickness of the IOBL differs significantly between ocean models and with model resolution. Furthermore, ocean models predict ice-shelf melting using parameterizations that neglect the buoyancy of the IOBL. This model deficiency likely biases turbulent fluxes through the IOBL and ocean cavity circulation, which is primarily driven by the buoyant flow of water freshened by ice-shelf melting. A new parameterization of ice-shelf melting that accounts for boundary layer dynamics is needed to achieve a more physically-based, accurate coupling of ice sheets and oceans in climate models.

% Overview of IOBL turbulence lit
IOBLs present unique conditions in the global ocean, involving a stabilizing flux from phase change (ice-melting) and a boundary layer that is positively buoyant against a sloping boundary. The IOBL has been most extensively explored at small scales, laboratory experiments and Direct Numerical Simulations, with some recent numerical studies addressing larger scales \cite{mondal_ablation_2019, vreugdenhil_stratification_2019}. However, the melt parameterizations commonly in use today still derive from sea-ice conditions (i.e., in the absence of a slope) \cite{mcphee_dynamics_1987}. Thus a knowledge gap exists in bridging these dynamics across the IOBL and characterizing the structure of buoyant plumes. Recently, this has been addressed through non-turbulence-resolving 2-d or 2.5-d models \cite{jenkins_simple_2016, jenkins_shear_2021, cheng_modeling_2020}. Notably, \citeA{jenkins_shear_2021} found a significant discrepancy between the commonly-used K-Profile Parameterization (KPP) and a higher-order turbulence closure scheme for vertical mixing in IOBL settings, suggesting that the application of KPP in ocean models may be inappropriate in ice-shelf cavities. 
There is a rich literature on stably-stratified boundary layers (typically under constant stabilizing flux boundary conditions), but the dependence of heat, salt and momentum fluxes on stratification remains a difficult problem, especially for strongly stratified regimes \cite{zonta_stably_2018}.

% Overview of this study
In this study, we model turbulent heat, salt, and momentum fluxes through the IOBL using Large-Eddy Simulation (LES). Whereas ocean models typically used to model sub-ice-shelf circulation cannot capture the relevant turbulent scales for boundary layer dynamics, LES captures the dominant energy-containing scales of turbulence and represents smaller, unresolved scales with varying degrees of complexity. An effective parameterization of ice-shelf melting is likely to rest on an understanding of how turbulent mixing in the IOBL depends on stratification and shear forcing. We vary far-field ocean temperature and ice-shelf slope between model runs and characterize turbulent fluxes, IOBL thickness, and ice-shelf melt rates. In this study we focus on the high-shear, low thermal driving regime. The target of previous LES studies has been on low-shear settings \cite{vreugdenhil_stratification_2019}.

\section{Methods}

@@

The model enforces incompressibility through a predictor-corrector method to the pressure perturbation, $\pi^*$, at each time step. Details can be found in \cite{maronga_parallelized_2015}. The momentum terms on the right hand side of Equation \ref{eq:uprog} are, in order: advection, Coriolis forcing, imposed geostrophic flow, buoyancy forcing, a correction for divergence in the flow (imposing incompressibility), and sub-grid scale momentum flux. Double primes denote sub-grid fluctuations.

@@

The terms on the right-hand side of the heat and salt conservation equations (Equations \ref{eq:ptprog} and \ref{eq:saprog}) are advection by the resolved flow, advection by the sub-grid scale fluctuations, and the source and sink terms  $\Phi_{\theta,S}$. The source and sink terms are zero in our simulations because we treat heat and salt fluxes due to melting at the boundaries through the sub-grid vertical flux term, as discussed in Section \ref{meth:tcm}.

@@

The turbulence closure scheme is Anisotropic Minimum Dissipation \cite{rozema_minimum-dissipation_2015} employing the extension to scalars introduced by \citeA{abkar_minimum-dissipation_2016}. We validated our implementation against the stable atmospheric boundary layer test case published in \citeA{abkar_large-eddy_2017}. Typical sub-grid scale momentum and scalar diffusivities in our sub-ice-shelf simulations of stratified turbulence range from 10\textsuperscript{-5} to 10\textsuperscript{-4} m\textsuperscript{2} s\textsuperscript{-1} in the upper two-thirds of the domain where damping is not applied (damping is discussed in Section \ref{meth:setup}.

Sources and sinks of momentum, heat and salt due to interactions between the flow and the ice base are all parameterized as sub-grid fluxes at the boundary. The resolved vertical fluxes at the top layer of the model go to zero as $w$ goes to zero according to the no-flux boundary condition and the sub-grid fluxes are determined using the scheme described below in place of the AMD scheme. PALM uses an Arakawa C-grid such that the ice boundary ($z=0$) is located at the edge of the top cell where the vertical velocity component resides and scalars and horizontal velocity components are located at mid-depth of the top cell ($z=-\frac{1}{2}\Delta_z$). We denote the interface location with subscript $b$ and the middle of the first cell below the interface with subscript $\frac{1}{2}$.

Momentum fluxes are parameterized according to law of the wall following a linear stability function as in \citeA{vreugdenhil_stratification_2019}:
%
\begin{equation} \label{eq:tau}
    \overline{w''u_i''}_b = u_* \frac{ \kappa \left[u_i\left(z_{\frac{1}{2}}\right) - u_b \right]}{\textrm{ln}\left(z_{\frac{1}{2}}/z_0\right)-\Psi_m\left(\zeta_{\frac{1}{2}}\right)}
\end{equation}
%
for the horizontal velocity components (i.e., $i=1,2$). $\zeta_k = z_k/L_O$ is the depth from the boundary scaled by the Monin-Obukhov length, $u_*$ is the friction velocity, and $z_0$ is the roughness length. The Monin-Obukhov length $L_O$, computed following \citeA{mcphee_dynamics_1987}, is a function of both $u_*$ and the melt rate. The stability function $\Psi_m$ is linear with the scaled depth:
\begin{equation}\label{eq:psi}
    \Psi_m(\zeta) = 1 + \beta_m \zeta
\end{equation}
This parameterization assumes that Coriolis forces are not dominant in the momentum balance.

Similarly the sub-grid fluxes of scalars are parameterized by
\begin{equation} \label{eq:melt_flux}
    \overline{w''\theta''}_b = \Gamma_{\theta} u_* \left(\theta_{\frac{1}{2}} - \theta_b\right),\: 
    \overline{w''S''}_b = \Gamma_S u_* \left(S_{\frac{1}{2}} - S_b\right)
\end{equation}
%
where the exchange coefficients are defined following \cite{mcphee_dynamics_1987}:
\begin{equation} \label{eq:gamma}
    \Gamma_{\theta}^{-1} = \left(\Gamma_{\theta,turb} + \Gamma_{\theta,mol}\right)^{-1},\:
    \Gamma_S^{-1} = \left(\Gamma_{S,turb} + \Gamma_{S,mol}\right)^{-1}
\end{equation}
%
The turbulent flux component follows a shape function consistent with the parameterization of momentum fluxes in Equation \ref{eq:tau}
\begin{equation} \label{eq:gamma_turb}
    \Gamma_{\theta,turb} = \kappa^{-1} \left[\textrm{ln}\left(z_{\frac{1}{2}}/z_0\right) - \Psi_{\theta}\left(\zeta_{\frac{1}{2}}\right)\right], \:
    \Gamma_{S,turb} = \kappa^{-1} \left[\textrm{ln}\left(z_{\frac{1}{2}}/z_0\right) - \Psi_S\left(\zeta_{\frac{1}{2}}\right)\right]    % + \Psi(z_0/L_O)) % correction for roughness length
\end{equation}
with analogous stability functions to momentum, Equation \ref{eq:psi}
\begin{equation}
    \Psi_{\theta}(\zeta) = 1 + \beta_{\theta}\zeta, \:
    \Psi_{S}(\zeta) = 1 + \beta_{S}\zeta
%    \Psi_{\theta,S}(\zeta) = 1 + \beta_{\theta,S} \zeta
\end{equation}
The coefficients of these stability functions are chosen as in \cite{zhou_self-similar_2017} and are given in Table \ref{table:var}. The molecular flux component follows \citeA{mcphee_dynamics_1987}:
\begin{equation}
    \Gamma_{\theta,mol} = 12.5 \textrm{Pr}^{2/3} - 6, \:
    \Gamma_{S,mol} = 12.5 \textrm{Sc}^{2/3} - 6
\end{equation}
where Pr is the Prandtl number and Sc is the Schmidt number. If a portion of the ice face experiences freezing, $\Gamma_\theta$ and $\Gamma_S$ are set to a higher value indicating destabilizing fluxes ($\Gamma_{f}$, Table \ref{table:var}).

@@

We present two sets of simulations which have a base case in common. The base case has low far-field thermal driving of 0.1$^{\circ}$C and a relatively high slope for Antarctic ice shelves of 1$^{\circ}$ and vigorous far-field inertial oscillations of 20 cm s\textsuperscript{-1} generated by a 0.03 Pa m\textsuperscript{-1} pressure gradient. These conditions were chosen to favor an energetic regime, for reasons discussed in Section \ref{disc:dyn}. To examine the relationship between thermal driving and melt rate, as well as turbulent flow characteristics, we vary the far-field thermal driving from 0.1$^{\circ}$C to 0.5$^{\circ}$C in the first set of simulations. In the second set of simulations, we reduce the slope from 1$^{\circ}$ to 0.01$^{\circ}$. The ice base always slopes to the east in the positive x-dimension of our domain, thus $\beta = 0$. The far-field salinity for all runs is 35 g kg\textsuperscript{-1}. Initially, both temperature and salinity increase with depth over the upper two-thirds of the domain. The background salinity stratification dominates the density stratification, with an inverse stability ratio $R_\rho^*$ of 20.

The simulation duration is $\sim$50 h, corresponding to $\sim$4 inertial periods of 13 h. By this point, the time-mean kinetic energy of the flow has reached steady-state. However, the turbulent intensity for all cases continues to decline, with more pronounced TKE loss at lower slopes (Figure \ref{fig:timeseries}a,d). Many of the results presented are averaged over the last simulated inertial period and only the depths outside the sponge layer are shown.

@@

\section{Results}

% Description of time evolution
Melt rates decline over the course of the simulations (Figure \ref{fig:timeseries}), preventing the identification of steady-state melt rates for most runs. This decline is in response to the concomitant increase in stratification during the course of the simulations, which decreases vertical heat fluxes by reducing vertical turbulent fluctuations. We do not continue our simulations beyond 4 inertial periods in the hope of reaching steady-state conditions because the increase in stratification reduces the turbulent length scales and necessitates higher model resolution than we have used here.

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dT_t.png}
    \end{minipage}
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dslope_t.png}
    \end{minipage}
    \caption{Time evolution of (a,d) domain-averaged turbulent kinetic energy (TKE), (b,e) melt rate and (c,f) friction velocity for (a-c) thermal driving simulations and (d-f) variable slope simulations. The black curve representing the same simulation in all panels.}
    \label{fig:timeseries}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip, width=\textwidth]{Figures/pt_cmp_dT_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip, width=\textwidth]{Figures/sa_cmp_dT_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip, width=\textwidth]{Figures/velocity_cmp_dT_43h_tav13h_z_profile.png}
    \end{minipage}
    \caption{Depth profiles of simulated properties as they vary with thermal driving, averaged over the last inertial period. (a) temperature relative to far-field temperature, (b) salinity relative to far-field salinity, and (c) velocity components. For panel (c), solid lines are in the positive y direction, dashed in positive x-direction (up-slope).}
    \label{fig:dT_profiles}
    % change colorbar label to \theta_\infty
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0cm 0 0},clip, width=\textwidth]{Figures/pt_cmp_dslope_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0cm 0 0},clip, width=\textwidth]{Figures/sa_cmp_dslope_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0cm 0 0},clip, width=\textwidth]{Figures/velocity_cmp_dslope_43h_tav13h_z_profile.png}
    \end{minipage}
    \caption{Depth profiles of simulated properties as they vary with slope, averaged over the last inertial period. (a) temperature relative to far-field temperature (b) salinity relative to far-field salinity (c) velocity components. For panel (c), solid lines are in the positive y direction, dashed in positive x-direction (up-slope). The black curve is common to both sets of simulations (see Figure \ref{fig:dT_profiles}).}
    \label{fig:dslope_profiles}
\end{figure}

% Description of velocity profiles
Ekman rotation near the boundary can be seen in all simulations (Figs. \ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c), but for more strongly sloped runs, buoyancy plays an increasingly strong role in driving the mean flow near the boundary. This effect can be seen most clearly by comparing the up-slope component of the flow within several meters of the ice-ocean interface across the slope-varying simulations (Figure \ref{fig:dslope_profiles}c). These mean buoyancy effects increase flow velocities on the order of a few cm s\textsuperscript{-1} as slope varies from 0.01$^{\circ}$ to 1.0$^{\circ}$ (Figure \ref{fig:dslope_profiles}c). Flow velocities near the boundary increase on the order of 10 cm s\textsuperscript{-1} as far-field thermal driving increases from 0.1 to 0.45$^{\circ}$C at 1$^{\circ}$ slope (Figure \ref{fig:dT_profiles}c). This velocity increase is attributed to changes in the magnitude of buoyancy forcing at the boundary, related to differences in melt rates.

>>>>

\documentclass[draft]{agujournal2019}
\usepackage{url} 
\usepackage{lineno}
\usepackage{graphicx}
%\usepackage{xr}
\usepackage{float}
%\externaldocument{A_suppl_fig}
\usepackage{amsmath}
\linenumbers
%\draftfalse
\journalname{JGR: Oceans}

@@

\textbf{Plain language summary}\\
Ice-shelf melting from below is controlled by the rate at which the ocean can transport heat vertically. Little is known about the small-scale mixing that carries this heat close to the ice base, where melting itself impedes mixing by forming a density barrier. This study uses ocean modeling at ultra-high resolution to study these mixing processes. It offers some insights into the relationship between ice-shelf melting and ocean temperature far from the ice base, which may help us project how fast ice will melt when ocean waters entering the cavity warm. This study adds to a growing body of research that indicates coarse resolution ocean models used for such projections may have to adjust how they handle ocean mixing to improve the fidelity of ice-shelf melting.

@@

% Motivation - SLR
The largest source of uncertainty in future sea level rise is the potential loss of ice from the Antarctic Ice Sheet \cite{ipcc_climate_2014}. The rate of grounded ice loss is highly sensitive to the melting of ice shelves, which drain over 80\% of Antarctica’s grounded ice \cite{reese_far_2018, rignot_ice-shelf_2013}. The Ice-shelf/Ocean Boundary Layer (IOBL) controls ice-shelf melting by regulating oceanic heat and salt fluxes to the ice shelf base. Thus, accurate predictions of ice-shelf melting depend on representing the turbulent dynamics of the IOBL. This representation is also critical for evaluating the sensitivity of the coupled land ice-ocean system to changes in ocean conditions. Of particular concern is the sensitivity of ice-shelf melting to increasing seawater temperature at depth, a trend observed along a wide swath of the West Antarctic coastline and a potential trigger for West Antarctic Ice Sheet collapse \cite{ruan_ice-shelf_2021, schmidtko_multidecadal_2014, purkey_unabated_2018, wahlin_pathways_2021}.

% Shortcomings of ocean-model representations of IOBL
One indication that ocean models do not capture turbulent dynamics is that the simulated thermal driving in the IOBL, the difference between ocean temperature and the local freezing point, and consequently the simulated melt rate differ significantly between ocean models and as resolution is varied within a model \cite{gwyther_cold_2020}. Furthermore, ocean models predict ice-shelf melting using parameterizations that neglect the lateral buoyancy gradient of the IOBL and often the vertical buoyancy gradient as well, which should reduce the efficiency of vertical mixing near the boundary. This model deficiency likely biases turbulent fluxes through the IOBL and ocean cavity circulation, which is driven in large part by the buoyant flow of water freshened by ice-shelf melting, overturning circulation also known as the ``ice pump'' \cite{webber_impact_2018}. A new parameterization of ice-shelf melting that accounts for boundary layer dynamics is needed to achieve a more physically based, accurate coupling of ice sheets and oceans in climate models \cite{gwyther_cold_2020, dinniman_modeling_2016, naughten_future_2018, edwards_projected_2021}.

% Overview of IOBL turbulence lit
IOBLs present unique conditions in the global ocean, involving a stabilizing flux from phase change (ice-melting) and a boundary layer that is positively buoyant against a sloping boundary. There is a rich literature on stably-stratified boundary layers (typically under constant stabilizing flux boundary conditions), but the dependence of heat, salt and momentum fluxes on stratification remains a difficult problem, especially for strongly stratified regimes \cite{zonta_stably_2018}. The IOBL has been most extensively explored at small scales through laboratory experiments and Direct Numerical Simulations, with some recent numerical studies addressing intermediate scales \cite{mondal_ablation_2019, vreugdenhil_stratification_2019, middleton_numerical_2021}. However, this body of work has not yet matured to setting a new standard for ice-shelf melt parameterization in ocean models. The commonly-used ice-shelf melt parameterization today still derives from sea-ice conditions (i.e., in the absence of a slope) with tuning of some parameters for ice-shelf settings \cite{holland_modeling_1999, jenkins_observation_2010, mcphee_dynamics_1987}. Thus a knowledge gap exists in bridging IOBL dynamics across scales and characterizing the structure of buoyant plumes. Recently, this has been addressed through non-turbulence-resolving 2-d or 2.5-d models \cite{jenkins_simple_2016, jenkins_shear_2021, cheng_modeling_2020}. Notably, \citeA{jenkins_shear_2021} found that vertical mixing in IOBL settings was represented poorly by the commonly-used K-Profile Parameterization (KPP) in comparison to a higher-order turbulence closure scheme, suggesting that the application of KPP in ocean models may be inappropriate in ice-shelf cavities.

% Overview of this study
In this study, we model turbulent heat, salt, and momentum fluxes through the IOBL using Large-Eddy Simulation (LES). Whereas ocean models typically used to model sub-ice-shelf circulation cannot capture the relevant turbulent scales for boundary layer dynamics, LES captures the dominant energy-containing scales of turbulence and represents smaller, unresolved scales with varying degrees of complexity. An effective parameterization of ice-shelf melting is likely to rest on an understanding of how turbulent mixing in the IOBL depends on stratification and shear forcing. We vary far-field ocean temperature and ice-shelf slope between model runs and characterize turbulent fluxes and ice-shelf melt rates. In this study we focus on the high-shear, low thermal driving regime. The target of previous LES studies has been on low-shear settings \cite{vreugdenhil_stratification_2019, middleton_numerical_2021}. Section \ref{meth} describes the LES model, its turbulence closure scheme, and the set-up of our simulations. Section \ref{results} presents the results of our simulations, with a focus on comparisons across the range of thermal driving and slope given. The discussion is given in two parts: Section \ref{disc:dyn} contextualizes the features of our simulated IOBL turbulence and discusses limitations of this study, and Section \ref{disc:prm} focuses on how this study informs parameterizations of ice-shelf melting and IOBL turbulence. We provide some closing thoughts in Section \ref{conclusion}.

\section{Methods}\label{meth}

@@

Detail on the solution of these equations can be found in \cite{maronga_parallelized_2015}. The momentum terms on the right hand side of Equation \ref{eq:uprog} are, in order: advection, Coriolis forcing, imposed geostrophic flow, buoyancy forcing, a correction for divergence in the flow using the pressure perturbation $\pi^*$ (imposing incompressibility), and sub-grid scale momentum flux. Double primes denote sub-grid fluctuations. All prognostic variables are considered filtered at the grid scale. This is typically denoted by the overbar, which we omit except for the sub-grid turbulent flux terms to emphasize that we only represent averaged effects.

@@

The terms on the right-hand side of the heat and salt conservation equations (Equations \ref{eq:ptprog} and \ref{eq:saprog}) are advection by the resolved flow, turbulent transport by the sub-grid scale fluctuations, and the source and sink terms  $\Phi_{\theta,S}$. The source and sink terms are zero in our simulations because we treat heat and salt fluxes due to melting at the boundaries through the sub-grid vertical flux term, as discussed in Section \ref{meth:tcm}.

@@

The turbulence closure scheme is Anisotropic Minimum Dissipation \cite<AMD;>{rozema_minimum-dissipation_2015} employing the extension to scalars introduced by \citeA{abkar_minimum-dissipation_2016}. We validated our implementation against the stable atmospheric boundary layer test case published in \citeA{abkar_large-eddy_2017}. Typical sub-grid scale momentum and scalar diffusivities in our sub-ice-shelf simulations of stratified turbulence range from 10\textsuperscript{-5} to 10\textsuperscript{-4} m\textsuperscript{2} s\textsuperscript{-1} in the upper two-thirds of the domain where damping is not applied (the damping methodology is discussed in Section \ref{meth:setup}.

Sources and sinks of momentum, heat and salt due to interactions between the flow and the ice base are all parameterized as sub-grid fluxes at the boundary. The resolved vertical fluxes at the top layer of the model go to zero as $w$ goes to zero according to the no-flux boundary condition and the sub-grid fluxes are determined using the scheme described below in place of the AMD scheme. PALM is vertically discretized such that the ice boundary ($z=0$) is located at the edge of the top cell where the vertical velocity component resides and scalars and horizontal velocity components are located at mid-depth of the top cell ($z=-\frac{1}{2}\Delta_z$). We denote the interface location with subscript $b$ and the middle of the first cell below the interface with subscript $\frac{1}{2}$.

Subgrid momentum fluxes are parameterized according to law of the wall following a linear stability function for stabilizing buoyancy forcing as in \citeA{vreugdenhil_stratification_2019}:
%
\begin{equation} \label{eq:tau}
    \overline{w''u_i''}_b = u_* \frac{ \kappa \left[u_i\left(z_{\frac{1}{2}}\right) - u_b \right]}{\textrm{ln}\left(z_{\frac{1}{2}}/z_0\right)-\Psi_m\left(\zeta_{\frac{1}{2}}\right)}
\end{equation}
%
for the horizontal velocity components (i.e., $i=1,2$). $\zeta_k = z_k/L_O$ is the depth from the boundary scaled by the Monin-Obukhov length, $u_*$ is the friction velocity, and $z_0$ is the roughness length. The Monin-Obukhov length $L_O$, computed following \citeA{mcphee_dynamics_1987}, is a function of both $u_*$ and the melt rate. The stability function $\Psi_m$ is linear with the scaled depth:
\begin{equation}\label{eq:psi}
    \Psi_m(\zeta) = 1 + \beta_m \zeta
\end{equation}
This parameterization assumes that Coriolis forces are not dominant in the momentum balance.

Similarly the sub-grid fluxes of scalars are parameterized by
\begin{equation} \label{eq:melt_flux}
    \overline{w''\theta''}_b = \Gamma_{\theta} u_* \left(\theta_{\frac{1}{2}} - \theta_b\right),\: 
    \overline{w''S''}_b = \Gamma_S u_* \left(S_{\frac{1}{2}} - S_b\right)
\end{equation}
%
where the exchange coefficients are defined following \cite{mcphee_dynamics_1987}:
\begin{equation} \label{eq:gamma}
    \Gamma_{\theta}^{-1} = \left(\Gamma_{\theta,turb} + \Gamma_{\theta,mol}\right)^{-1},\:
    \Gamma_S^{-1} = \left(\Gamma_{S,turb} + \Gamma_{S,mol}\right)^{-1}
\end{equation}
%
The turbulent flux component follows a shape function consistent with the parameterization of momentum fluxes in Equation \ref{eq:tau}
\begin{equation} \label{eq:gamma_turb}
    \Gamma_{\theta,turb} = \kappa^{-1} \left[\textrm{ln}\left(z_{\frac{1}{2}}/z_0\right) - \Psi_{\theta}\left(\zeta_{\frac{1}{2}}\right)\right], \:
    \Gamma_{S,turb} = \kappa^{-1} \left[\textrm{ln}\left(z_{\frac{1}{2}}/z_0\right) - \Psi_S\left(\zeta_{\frac{1}{2}}\right)\right]    % + \Psi(z_0/L_O)) % correction for roughness length
\end{equation}
with analogous stability functions to momentum, Equation \ref{eq:psi}
\begin{equation}
    \Psi_{\theta}(\zeta) = 1 + \beta_{\theta}\zeta, \:
    \Psi_{S}(\zeta) = 1 + \beta_{S}\zeta
%    \Psi_{\theta,S}(\zeta) = 1 + \beta_{\theta,S} \zeta
\end{equation}
The coefficients of these stability functions are chosen as in \citeA{zhou_self-similar_2017} and are given in Table \ref{table:var}. The molecular flux component follows \citeA{mcphee_dynamics_1987}:
\begin{equation}
    \Gamma_{\theta,mol} = 12.5 \textrm{Pr}^{2/3} - 6, \:
    \Gamma_{S,mol} = 12.5 \textrm{Sc}^{2/3} - 6
\end{equation}
where Pr is the Prandtl number and Sc is the Schmidt number. If a portion of the ice face experiences freezing, $\Gamma_\theta$ and $\Gamma_S$ are set to a higher value indicating destabilizing fluxes ($\Gamma_{f}$, Table \ref{table:var}).

@@

We present two sets of simulations which have a base case in common. The base case has low far-field thermal driving of 0.15$^{\circ}$C and a relatively high slope for Antarctic ice shelves of 1$^{\circ}$ and vigorous far-field inertial oscillations of 20 cm s\textsuperscript{-1} generated by a 0.03 Pa m\textsuperscript{-1} pressure gradient. These conditions were chosen to favor an energetic regime, for reasons discussed in Section \ref{disc:dyn}. To examine the relationship between thermal driving and melt rate, as well as turbulent flow characteristics, we vary the far-field thermal driving from 0.15$^{\circ}$C to 0.6$^{\circ}$C in the first set of simulations. Far-field thermal driving $\Delta \theta_\infty$ is defined as $\theta_{\infty} - \theta_f(S_{\infty})$. In the second set of simulations, we reduce the slope from 1$^{\circ}$ to 0.01$^{\circ}$. The ice base always slopes to the east in the positive x-dimension of our domain, thus $\beta = 0$. The far-field salinity for all runs is 35 g kg\textsuperscript{-1}. Initially, both temperature and salinity increase with depth over the upper two-thirds of the domain. The background salinity stratification dominates the density stratification, with an inverse stability ratio $R_\rho^*$ of 20.

The simulation duration is $\sim$50 h, corresponding to $\sim$4 inertial periods of 13 h. By the end of our simulations, the time-mean kinetic energy of the flow has reached steady-state. However, the turbulent intensity for all cases continues to decline, with more pronounced turbulence kinetic energy (TKE) loss at lower slopes (Figure \ref{fig:timeseries}a,d). Many of the results presented are averaged over the last simulated inertial period and only the depths outside the sponge layer are shown.

We compute an effective thermal exchange coefficient that differs from that employed by the sub-grid scheme to represent the efficiency of heat exchange that may need to be represented to produce accurate melt rates in a model that does not have the vertical resolution or sophisticated turbulence closure that we employ here. This derived thermal exchange coefficient, $\Gamma_{T,der}$, can be computed using Equations~\ref{eq:melt_flux} and \ref{eq:melt_1} from the simulated melt rate and ocean properties at any depth rather than $z_{\frac{1}{2}}$. We substitute the friction velocity computed by Equation~\ref{eq:tau} with one computed using a quadratic drag law from the velocity at $z < z_{\frac{1}{2}}$ and the applied drag coefficient, consistent with modeled drag implementations.

@@

\section{Results}\label{results}

% Description of time evolution
Melt rates decline over the course of the simulations (Figure \ref{fig:timeseries}), preventing the identification of steady-state melt rates for most runs. This decline is in response to the concomitant increase in stratification during the course of the simulations, which decreases vertical heat fluxes by reducing vertical turbulent fluctuations. We do not continue our simulations beyond 4 inertial periods in the hope of reaching steady-state conditions because the increase in stratification reduces the turbulent length scales and necessitates higher model resolution than we have used here. Our simulations generally show resolved turbulent fluxes exceeding the subgrid turbulent fluxes, but over the last inertial period the subgrid turbulent fluxes dominate within several meters of the ice base where the stratification is strongest (Figure S1).

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dT_t.png}
    \end{minipage}
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dslope_t.png}
    \end{minipage}
    \caption{Time evolution of (a,d) domain-averaged turbulence kinetic energy (TKE), (b,e) melt rate and (c,f) friction velocity for (a-c) thermal driving simulations and (d-f) variable slope simulations. The black curve represents the same simulation in all panels.}
    % Consider shading the analysis window
    % TODO add panel labels
    \label{fig:timeseries}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip, width=\textwidth]{Figures/pt_cmp_dT_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip, width=\textwidth]{Figures/sa_cmp_dT_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip, width=\textwidth]{Figures/velocity_cmp_dT_43h_tav13h_z_profile.png}
    \end{minipage}
    \caption{Depth profiles of simulated properties as they vary with thermal driving, averaged over the last inertial period. (a) temperature relative to far-field temperature, (b) salinity relative to far-field salinity, and (c) velocity components. For panel (c), solid lines are in the y direction (positive north), dashed in x-direction (positive east, up-slope).}
    \label{fig:dT_profiles}
    % TODO add panel labels
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0cm 0 0},clip, width=\textwidth]{Figures/pt_cmp_dslope_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0cm 0 0},clip, width=\textwidth]{Figures/sa_cmp_dslope_43h_tav13h_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0cm 0 0},clip, width=\textwidth]{Figures/velocity_cmp_dslope_43h_tav13h_z_profile.png}
    \end{minipage}
    \caption{Depth profiles of simulated properties as they vary with slope, averaged over the last inertial period. (a) temperature relative to far-field temperature (b) salinity relative to far-field salinity (c) velocity components. For panel (c), solid lines are in the positive y direction, dashed in positive x-direction (up-slope). The black curve is common to both sets of simulations (see Figure \ref{fig:dT_profiles}).}
    \label{fig:dslope_profiles}
    % TODO add panel labels
\end{figure}

% Description of velocity profiles
Ekman rotation near the boundary can be seen in all simulations (Figs. \ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c), but for more strongly sloped runs, buoyancy plays an increasingly strong role in driving the mean flow near the boundary. This effect can be seen most clearly by comparing the up-slope component of the flow within several meters of the ice-ocean interface across the slope-varying simulations (Figure \ref{fig:dslope_profiles}c). These mean buoyancy effects increase flow velocities on the order of a few cm s\textsuperscript{-1} as slope varies from 0.01$^{\circ}$ to 1$^{\circ}$ (Figure \ref{fig:dslope_profiles}c). Flow velocities near the boundary increase on the order of 10 cm s\textsuperscript{-1} as far-field thermal driving increases from 0.5 to 0.6$^{\circ}$C at 1$^{\circ}$ slope (Figure \ref{fig:dT_profiles}c). This velocity increase is attributed to changes in the magnitude of buoyancy forcing at the boundary, related to differences in melt rates. The time-averaged far-field velocity shown in Figures~\ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c integrates inertial oscillations of the geostrophic velocity at 20 cm s\textsuperscript{-1}.
